Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-02-05 vesicle_code/docs/adhesion/adhesion.tex
read entire paper and made edits/comments through out

The dynamics of weakly adhesive vesicles under various flowing conditions is investigated numerically using a boundary integral formulation with
adaptive time-marching that facilitates efficient and accurate computations of vesicle hydrodynamics under adhesion. In a quiescent flow, two nearby vesicles 
move slowly towards each other under the adhesive potential, pushing out fluid in the gap to  
form a doublet at equilibrium. A lubrication analysis on such draining of the
thin film gives the dependencies of draining time on adhesion strength and separation distance that are in good agreement with results from
boundary integral simulations.
%
In a microfluid trap where the stagnation of an extensional flow  is placed in the middle of a vesicle doublet through an active control loop,
novel dynamics of a vesicle doublet is observed. Results from boundary integral simulations show that there exists a critical extensional flow rate above which the weak adhesive interaction is overcome by the hydrodynamic diverging stream, thus providing a simple method to measure the adhesion strength between two vesicle membranes. 
%
In a planar shear flow, numerical simulations reveal that a vesicle
doublet may form as they approach each other in a planar shear flow
provided that the adhesion strength is sufficient large at a given
vesicle reduced area. Once a doublet the effective shear viscosity of a
dilute suspension of vesicle doublets is found to be a function of both
reduced area and the adhesion strength.  Results from these numerical
studies and analysis shed light on the hydrodynamic effects of adhesive
interactions between vesicles in a viscous fluid.
%\begin{description}
%\item[Usage]
%Secondary publications and information retrieval purposes.
%\item[PACS numbers]
%May be entered using the \verb+\pacs{#1}+ command.
%\item[Structure]
%You may use the \texttt{description} environment to structure your abstract;
%use the optional argument of the \verb+\item+ command to give the category of each item. 
%\end{description}
\end{abstract}

@@

Hydrodynamics of a single vesicle in Stokes flow has been extensively
investigated. In a planar shear flow, the vesicle hydrodynamics is
characterized by the reduced volume (reduced area in two dimensions),
viscosity contrast between interior and exterior fluids, and shear rate
of the imposed far-field fluid flow. In addition, a vesicle with a rigid
particle inside is also investigated as a biological mimic of a
eukaryotic cell with a nucleus that occupies nearly half of the
intracellular volume~\cite{Veerapaneni2011_PRL}. Small-deformation
analysis shows that a vesicle tank-treads in a planar shear flow for low
viscosity contrast and shear rate. At high viscosity contrast the
tank-tread (TT) dynamics transitions to tumbling (TB)
dynamics~\cite{Misbah2006_PRL, Vlahovska2007_PRE}, and this leads to a
transition in effective shear viscosity in the vesicle
suspension~\cite{Misbah2006_PRL,Vitkova2008_BJ} that is also validated
by direct numerical simulations~\cite{GhigliottiBibenMisbah2010_JFM} and
experiments~\cite{KantslerSegreSteinberg2008_EPL,
ZabuskySegreDeschamps2011_PoF}.  Between TT and TB vesicle dynamics, a
breathing (tremble) mode is also
observed~\cite{Misbah2006_PRL,KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2011_JFM, SpannZhaoShaqfeh2014_PoF} to affect the effective
shear viscosity in a different way.  In an extensional flow (planar or
uniaxial), vesicle shape dynamics depends sensitively on the vesicle
reduced volume and the elastic capillary
number~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}: Asymmetric shape
and oscillatory undulation of the vesicle membrane are two examples of
the complex vesicle hydrodynamics in an extensional flow.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in a shear flow}
\label{sec:sflow} 
We consider two vesicles suspended in the shear flow $\uu = \chi(y,0)$,
where $\chi$ is the shear rate.  The vesicles are placed on the $y$ axis
so that the background velocity drives the vesicles towards one another.
In the absence of adhesion, once the vesicles are sufficiently close,
they are deflected to opposite sides of the $y$ axis, pass one another,
and separate from one another.  However, in the presence of adhesion,
the vesicles can form a doublet for certain values of the adhesion
length scale $\delta$, Hamaker constant $\mathcal{H}$, shear rate
$\chi$, and reduced area $\Delta A$.  Figure~\ref{fig:doublet090} shows
snapshots two vesicles that have formed a doublet and the color coding
is the tension along the vesicle.  The two vesicles are
identical with reduced area is $\Delta A = 0.9$, the shear rate is $\chi
= 0.5$, the adhesion length scale is $\delta = 0.4$, and the Hamaker
constant is $\mathcal{H} = 0.7$.  Similar to the quiescent example, we
observe that the membrane tension in the contact region is negative in
the contact region indicating that the membrane is being compressed when
the adhesive force is strongest.

@@

In the left plot of Figure~\ref{fig:shearIntrinsicViscosity}, we compare
the intrinsic viscosity of a single tank treading vesicle to a doublet.
To validate our simulations, we superimpose (black marks) the intrinsic
viscosity calculated by Ghigliotti et
al.~\cite{GhigliottiBibenMisbah2010_JFM}.  The presence of the doublet
significantly increases the intrinsic viscosity at all the reduced
areas.  To further characterize the effect of adhesion, in the right
plot of Figure~\ref{fig:shearIntrinsicViscosity} we decompose the
intrinsic viscosity into the contributions from the bending and tension
(blue), and the contribution from the adhesion (red).  We also
superimpose twice the intrinsic viscosity of a single tank treading
vesicle (dashed curve) to demonstrate that the bending and tension of
the doublet behave similarly, but not identically, to a dilute
suspensions of non-adhering tank-treading vesicles.  We see that the
effect of the adhesion on the intrinsic viscosity is largest for
vesicles with small reduced areas.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.45\textwidth]{figs/shear2Ves_adR4em1adS7em1Chi5em1.pdf}
  \includegraphics[width=0.45\textwidth]{figs/doublet_decomp.pdf}
  \caption{\label{fig:shearIntrinsicViscosity} {\em Left}: The intrinsic
  viscosity of a tank treading vesicle (blue) and a doublet (red).  The
  shear rate is $\chi = 0.5$, the Hamaker constant is $\mathcal{H} =
  0.7$, and the adhesion length scale is $\delta = 0.4$, and the length
  of the vesicles are held fixed.  The black marks denote intrinsic
  viscosity values computed by Ghigliotti {\em et
  al.}~\cite{GhigliottiBibenMisbah2010_JFM} (cf.~Figure 5).  {\em
  Right}: The decomposition of the intrinsic viscosity into the
  contributions for the bending and tension (blue) of the vesicles, and
  the contribution from the vesicle adhesion (red).  Also included is
  twice the intrinsic viscosity of a single tank-treading vesicle
  (dashed black line).}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions\label{sec:conclusions}}
In this work we use boundary integral formulation with adaptive
time-stepping  to simulate hydrodynamics of two vesicles with adhesive
interactions.  In a quiescent flow,  two vesicles move towards each
other under a long-range attraction when they are in close vicinity of
each other.  We use a lubrication theory to estimate the time it takes
to reach the separation distance, and found the theoretical scaling to
be in good agreement with numerical results.  Once two membranes are
within separation distance $\delta$, the adhesive force turns repulsive
and the membranes flatten to form a contact region.  Our simulations
show that the membranes in the contact region are actually curved with
end points at the shortest distance.  Once a vesicle doublet forms, we
examine the dependence of membrane bending and adhesion energies on
reduced area, Hamakar constant and separation distance.

Next we conduct a numerical experiment where a vesicle doublet is placed
at the center of a fluid trap, which can be actively controlled in
microfluidic channel so that fluid trap center is effectively the
stagnation point of an extensional flow characterized by a extensional
flow rate.  At low flow rate, we find the doublet to rotate ninety
degrees to align with the flow such that the long axis of the doublet is
along the divergent axis and the convergent stream is pushing the two
vesicles together. As the flow rate increases the vesicle doublet
rotates less and we observe tank-treading like motion of both vesicles
in the doublet.  When the flow rate reaches above the critical value,
the diverging flow pulls the vesicles apart before the doublet has time
to rotate to remain intact.  These results indicate that it is possible
to use the fluid trap to separate a vesicle doublet under adhesion, and
thus provide a means to probe the adhesion strength between membranes.
For a pair of $\mu m$-sized vesicles with a bending rigidity of $\sim
10^{-19}$ $J$, an extensional flow rate of $0.5$ $s^{-1}$ is expected to
separate a vesicle doublet with reduced area of $\Delta A = 0.8$ and a
Hamakar constant $\mathcal{H}=0.7$, which corresponds to $\sim 1$ $\mu
J/m^2$.  These conditions are quite realizable in micrifulidic
experiments, and we hope that our simulations might inspire microfluidic
experiments.

We also examine how adhesive interaction may lead to formation of a
vesicle doublet dynamically in flowing conditions.  We simulate two
vesicles approaching each other in a planar shear flow, and examine how
their adhesive interactions lead to doublet formation.  Once a doublet
forms, the two vesicle membranes rotate around each other as they deform
dynamically. The usual tank-treading motion of a vesicle under shear
flow is not observed in each of the two vesicles. We also compute the
effective shear viscosity of a dilute suspension of vesicle doublets,
and found that membrane adhesion contribution increases with decreasing
reduced area because deflated vesicles (low reduced area) deform more
easily and more membranes in the contact region.

@@

Future work includes three dimensional simulations, viscosity contrast,
confined suspensions, and dense suspensions.

@@

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot01.pdf}
  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot02.pdf}
  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot03.pdf}
  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot04.pdf} \\
  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot05.pdf}
  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot06.pdf}
  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot07.pdf}
  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot08.pdf}
  \caption{\label{figs:relaxationManyVes} Five vesicles suspended inside
  a large vesicle.  The background flow is quiescent, the adhesion
  length scale is $\delta = 0.4$, and the Hamaker constant is
  $\mathcal{H} = 0.7$.  At the final snapshot, the suspension is near an
  equilibrium configuration.}
\end{figure}

>>>>

The dynamics of weakly adhesive vesicles under various flowing
conditions is investigated numerically using a high-order
adaptive-in-time boundary integral equation formulation. In a quiescent
flow, two vesicles move slowly towards each other under the adhesive
potential, pushing out fluid in the gap to form a doublet at
equilibrium. A lubrication analysis on such draining of the thin film
gives the dependencies of draining time on adhesion strength and
separation distance that are in good agreement with numerical results.
%
In a microfluid trap where the stagnation of an extensional flow is
placed in the middle of a vesicle doublet through an active control
loop, novel dynamics of a vesicle doublet are observed. Numerical
simulations show that there exists a critical extensional flow rate
above which the weak adhesive interaction is overcome by the
hydrodynamic diverging stream, thus providing a simple method to measure
the adhesion strength between two vesicle membranes. 
%
In a planar shear flow, numerical simulations reveal that a vesicle
doublet may form as they approach each other provided that the adhesion
strength is sufficient large at a given vesicle reduced area. Once a
doublet is formed, the effective shear viscosity of a dilute suspension
of vesicle doublets is found to be a function of both reduced area and
the adhesion strength.  Results from these numerical studies and
analysis shed light on the hydrodynamic effects of adhesive interactions
between vesicles in a viscous fluid.
%\begin{description}
%\item[Usage]
%Secondary publications and information retrieval purposes.
%\item[PACS numbers]
%May be entered using the \verb+\pacs{#1}+ command.
%\item[Structure]
%You may use the \texttt{description} environment to structure your abstract;
%use the optional argument of the \verb+\item+ command to give the category of each item. 
%\end{description}
\end{abstract}

@@

Hydrodynamics of a single vesicle in Stokes flow has been extensively
investigated. In a planar shear flow, the vesicle hydrodynamics is
characterized by the reduced volume (reduced area in two dimensions),
viscosity contrast between interior and exterior fluids, and shear rate
of the imposed far-field fluid flow. In addition, a vesicle with a rigid
particle inside is also investigated as a biological mimic of a
eukaryotic cell with a nucleus that occupies nearly half of the
intracellular volume~\cite{Veerapaneni2011_PRL}. Small-deformation
analysis shows that a vesicle tank-treads in a planar shear flow for low
viscosity contrast and shear rate. At high viscosity contrast the
tank-treading dynamics transitions to tumbling 
dynamics~\cite{Misbah2006_PRL, Vlahovska2007_PRE}, and this leads to a
transition in effective shear viscosity in the vesicle
suspension~\cite{Misbah2006_PRL,Vitkova2008_BJ} that is also validated
by direct numerical simulations~\cite{GhigliottiBibenMisbah2010_JFM} and
experiments~\cite{KantslerSegreSteinberg2008_EPL,
ZabuskySegreDeschamps2011_PoF}.  Between tank-treading and tumbling
vesicle dynamics, a breathing (tremble) mode is also
observed~\cite{Misbah2006_PRL,KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2011_JFM, SpannZhaoShaqfeh2014_PoF} to affect the effective
shear viscosity in a different way.  In an extensional flow (planar or
uniaxial), vesicle shape dynamics depends sensitively on the vesicle
reduced volume and the elastic capillary
number~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}: Asymmetric shape
and oscillatory undulation of the vesicle membrane are two examples of
the complex vesicle hydrodynamics in an extensional flow.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in a shear flow}
\label{sec:sflow} 
We consider two vesicles suspended in the shear flow $\uu = \chi(y,0)$,
where $\chi$ is the shear rate.  The vesicles are placed on the $y$ axis
so that the background velocity drives the vesicles towards one another.
In the absence of adhesion, once the vesicles are sufficiently close,
they are deflected to opposite sides of the $y$ axis, pass one another,
and separate from one another.  However, in the presence of adhesion,
the vesicles can form a doublet for certain values of the adhesion
length scale $\delta$, Hamaker constant $\mathcal{H}$, shear rate
$\chi$, and reduced area $\Delta A$.  Figure~\ref{fig:doublet090} shows
snapshots two vesicles that have formed a doublet and the color coding
is the tension along the vesicle.  The two vesicles are
identical with reduced area is $\Delta A = 0.9$, the shear rate is $\chi
= 0.5$, the adhesion length scale is $\delta = 0.4$, and the Hamaker
constant is $\mathcal{H} = 0.7$.  Similar to the quiescent example, we
observe that the membrane tension in the contact region is negative in
the contact region indicating that the membrane is being compressed when
the adhesive force is strongest.  Once the vesicles have formed a
doublet, the vesicles rotate in tandem rather than performing the usual
tank-treading motion.

@@

In the left plot of Figure~\ref{fig:shearIntrinsicViscosity}, we compare
the intrinsic viscosity of a single tank-treading vesicle to a doublet.
To validate our simulations, we superimpose (black marks) the intrinsic
viscosity calculated by Ghigliotti {\em et
al.}~\cite{GhigliottiBibenMisbah2010_JFM}.  The presence of the doublet
significantly increases the intrinsic viscosity at all the reduced
areas.  To further characterize the effect of adhesion, in the right
plot of Figure~\ref{fig:shearIntrinsicViscosity} we decompose the
intrinsic viscosity into the contributions from the bending and tension
(blue), and the contribution from the adhesion (red).  We also
superimpose twice the intrinsic viscosity of a single tank-treading
vesicle (dashed curve) to demonstrate that the bending and tension of
the doublet behave similarly, but not identically, to a dilute
suspensions of non-adhering tank-treading vesicles.  We see that the
effect of the adhesion on the intrinsic viscosity is largest for
vesicles with small reduced areas.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.45\textwidth]{figs/shear2Ves_adR4em1adS7em1Chi5em1.pdf}
  \includegraphics[width=0.45\textwidth]{figs/doublet_decomp.pdf}
  \caption{\label{fig:shearIntrinsicViscosity} {\em Left}: The intrinsic
  viscosity of a tank-treading vesicle (blue) and a doublet (red).  The
  shear rate is $\chi = 0.5$, the Hamaker constant is $\mathcal{H} =
  0.7$, and the adhesion length scale is $\delta = 0.4$, and the length
  of the vesicles are held fixed.  The black marks denote intrinsic
  viscosity values computed by Ghigliotti {\em et
  al.}~\cite{GhigliottiBibenMisbah2010_JFM} (cf.~Figure 5).  {\em
  Right}: The decomposition of the intrinsic viscosity into the
  contributions for the bending and tension (blue) of the vesicles, and
  the contribution from the vesicle adhesion (red).  Also included is
  twice the intrinsic viscosity of a single tank-treading vesicle
  (dashed black line).}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions\label{sec:conclusions}}
In this work we use boundary integral formulation with adaptive
time-stepping to simulate hydrodynamics of two vesicles with adhesive
interactions.  In a quiescent flow, two vesicles that are initially
sufficiently far apart move towards each other under a long-range
attraction.  We use a lubrication theory to estimate the time required
to reach the separation distance, and the theoretical scaling is in good
agreement with numerical results.  Once two membranes are within
separation distance $\delta$, the adhesive force turns repulsive and the
membranes flatten to form a contact region.  Our simulations show that
the membranes in the contact region are actually curved with end points
at the shortest distance.  Once a vesicle doublet forms, we examine the
dependence of membrane bending and adhesion energies on the reduced
area, the Hamaker constant, and the separation distance.

Next we conduct a numerical experiment where a vesicle doublet is placed
at the center of a fluid trap, which can be actively controlled in
microfluidic channel so that fluid trap center is effectively the
stagnation point of an extensional flow characterized by a extensional
flow rate.  At low flow rate, we find the doublet to rotate nearly ninety
degrees to align with the flow such that the long axis of the doublet is
along the divergent axis and the convergent stream is pushing the two
vesicles together. As the flow rate increases, the vesicle doublet
rotates less, and when the flow rate reaches above the critical value,
the diverging flow breaks the doublet structure by pulling the vesicles
apart.  These results indicate that it is possible to use the fluid trap
to separate a vesicle doublet under adhesion, and thus provide a means
to probe the adhesion strength between membranes.  For a pair of $\mu
m$-sized vesicles with a bending rigidity of $\sim 10^{-19}$ $J$, an
extensional flow rate of $0.5$ $s^{-1}$ is expected to separate a
vesicle doublet with reduced area of $\Delta A = 0.8$ and a Hamaker
constant $\mathcal{H}=0.7$, which corresponds to $\sim 1$ $\mu J/m^2$.
These conditions are quite realizable in microfluidic experiments, and
we hope that our simulations will inspire microfluidic experiments.

We also examine how adhesive interaction may lead to formation of a
vesicle doublet dynamically in flowing conditions.  We simulate two
vesicles approaching each other in a planar shear flow, and examine how
their adhesive interactions lead to doublet formation.  Once a doublet
forms, the two vesicle membranes rotate around each other as they deform
dynamically. The usual tank-treading motion of a vesicle under shear
flow is not observed in each of the two vesicles. We also compute the
effective shear viscosity of a dilute suspension of vesicle doublets,
and found that membrane adhesion contribution increases with decreasing
reduced area because deflated vesicles (low reduced area) deform more
easily and more membranes in the contact region.
\todo[inline]{I don't understand this last sentence}

@@

Future work includes three dimensional simulations, viscosity contrast,
dense suspensions, and confined suspensions.  In
Figure~\ref{fig:relaxationManyVes} illustrate snapshots of five vesicles
suspended inside of a larger vesicle.  There is no imposed flow, so the
dynamics are governed entirely by the adhesion, bending, and tension
forces.  In the absence of confinement, we would expect the vesicles to
align themselves as a rouleaux with the membranes forming a parachute,
male-female, or S-shape
configurations~\cite{FlormannAouane2017_SciReports}.  However, because
of the bounding vesicle and the dense packing, the interior vesicles
align themselves differently.  In addition, several parts of the
bounding membrane closely follow the interior membranes in order to
minimize the adhesion energy.  In
Figure~\ref{fig:relaxationManyVesAdhesionTension} we plot the tension
(left) and adhesion (right) of the final equilibrium configuration.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot01.pdf}
  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot02.pdf}
  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot03.pdf}
  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot04.pdf} \\
  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot05.pdf}
  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot06.pdf}
  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot07.pdf}
  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot08.pdf}
  \caption{\label{fig:relaxationManyVes} Five vesicles suspended inside
  a large vesicle.  The background flow is quiescent, the adhesion
  length scale is $\delta = 0.4$, and the Hamaker constant is
  $\mathcal{H} = 0.7$.  At the final snapshot, the suspension is near an
  equilibrium configuration.}
\end{figure}

@@

\begin{figure}[htp]
\todo[inline]{Figures of tension and adhesion of the final
configuration}
\caption{\label{fig:relaxationManyVesAdhesionTension} Something smart to
say ....}
\end{figure}
